---
title: "Inglis et al. (2024)"
format:
  html:
    df-print: kable
    code-fold: true
  markdown_github: 
    df-print: kable
---

```{r}
#| warning: false
library(tidyverse)
```


## Data

```{r}
ref2021 <- read_csv(here::here("inglis-et-al-2024/REF2021_UnitResults35.csv"), show_col_types = FALSE)
```

Topics:

```{r}
topic_names <- ref2021 %>% 
  select(matches("^T(\\d+)")) %>% 
  names()

topic_summary <- topic_names %>% 
  enframe() %>%  
  select(-name) %>% 
  separate_wider_delim(value, delim = " - ", names = c("code", "description"))

topic_summary
```

```{r}
ref2021_clean <- ref2021 %>%
  rename_with(.cols = matches("^T(\\d+)"),
              .fn = ~ str_extract(.x, "^(T(\\d+))")) %>%
  select(-starts_with("Log2"))
```


Here we reconstruct the transformed columns, where each of the "Tx" columns produces a corresponding "Log2(Tx)_Log2(T3)" column, with the values being $\log_2(Tx/T3)$.

```{r}
ref2021_Log2T3 <- ref2021_clean %>%
  mutate(across(
    matches("^T(\\d+)"),
    .fns = ~ log2(.x / T3),
    .names = "Log2({.col})_Log2(T3)"
  ))
```

## Regression

```{r}
ref2021_Log2T3_regression_data <- ref2021_Log2T3 %>% 
  select(OutputsGPA, starts_with("Log2")) %>% 
  select(-starts_with("Log2(T3)"))

ref2021_Log2T3_regression <-
  lm(OutputsGPA ~ ., data = ref2021_Log2T3_regression_data)

summary(ref2021_Log2T3_regression)
```

This replicates the $R^2$ value of `r summary(ref2021_Log2T3_regression)$r.squared`.

Regression coefficients:

```{r}
ref2021_Log2T3_regression_coefficients <-
  ref2021_Log2T3_regression$coefficients %>% 
  as_tibble(rownames = "predictor") %>% 
  rename(coefficient = value) %>% 
  arrange(-coefficient) %>% 
  mutate(topic = case_when(
    predictor == "(Intercept)" ~ "Intercept",
    .default = str_extract(predictor, "Log2\\((T\\d+)\\)", group = 1)
  )) %>% 
  left_join(topic_summary, by = c("topic" = "code")) %>% 
  mutate(description = if_else(topic == "Intercept", "Intercept", description)) %>% 
  select(topic, description, coefficient)

ref2021_Log2T3_regression_coefficients
```

This is missing the coefficient for T3. We can reconstruct that as the negative of the sum of the other topic coefficients.

```{r}
ref2021_Log2T3_regression_coefficients %>% 
  filter(topic != "Intercept") %>% 
  summarise(T3 = -sum(coefficient))
```

(This makes sense, based on equation (7) of Coenders, G., & Pawlowsky-Glahn, V. (2020). On interpretations of tests and effect sizes in regression models with a compositional predictor. _SORT, 44_, 201–220. <https://doi.org/10.2436/20.8080.02.100> )

## Why T3?

Do we get the same results if we use a different topic in the denominator instead of T3?

Let's try T1...

```{r}
ref2021_Log2T1 <- ref2021_clean %>%
  mutate(across(
    matches("^T(\\d+)"),
    .fns = ~ log2(.x / T1),
    .names = "Log2({.col})_Log2(T1)"
  ))

ref2021_Log2T1_regression_data <- ref2021_Log2T1 %>% 
  select(OutputsGPA, starts_with("Log2")) %>% 
  select(-starts_with("Log2(T1)"))

ref2021_Log2T1_regression <-
  lm(OutputsGPA ~ ., data = ref2021_Log2T1_regression_data)
```

The $R^2$ value is `r summary(ref2021_Log2T1_regression)$r.squared`.

Regression coefficients:

```{r}
ref2021_Log2T1_regression_coefficients <-
  ref2021_Log2T1_regression$coefficients %>% 
  as_tibble(rownames = "predictor") %>% 
  rename(coefficient = value) %>% 
  arrange(-coefficient) %>% 
  mutate(topic = case_when(
    predictor == "(Intercept)" ~ "Intercept",
    .default = str_extract(predictor, "Log2\\((T\\d+)\\)", group = 1)
  )) %>% 
  left_join(topic_summary, by = c("topic" = "code")) %>% 
  mutate(description = if_else(topic == "Intercept", "Intercept", description)) %>% 
  select(topic, description, coefficient)

ref2021_Log2T1_regression_coefficients

ref2021_Log2T1_regression_coefficients %>% 
  filter(topic != "Intercept") %>% 
  summarise(T1 = -sum(coefficient))
```


## Train and test

Here we try to fit the regression model using only 70 of the UoAs; for the remaining 13 UoAs, we then compare the predicted output GPA from the model with the actual GPA. We do this train/test split 1000 times, recording the $R^2$ value each time to get a sort of bootstrap estimate for the predictive value of the regression model. (Note that the results are sensitive to the size of the training set used, with 70 being a somewhat arbitrary choice.)

```{r}
set.seed(20241028)
all_dat <- ref2021_Log2T3_regression_data %>% 
  mutate(id = row_number())
train_size <- 70

rsquared_values <- c()

for (i in 1:1000) {
  train <- all_dat %>% slice_sample(n = train_size)
  test <- all_dat %>% anti_join(train, by = "id")
  
  regression <-
    lm(OutputsGPA ~ ., data = train %>% select(-id))
  
  predictions <- bind_cols(
    test %>% select(-starts_with("Log2")),
    predict(regression, newdata = test) %>% as_tibble()
  ) %>% 
    rename(predictedGPA = value)
  
  rsquared <- cor(predictions$OutputsGPA, predictions$predictedGPA)^2
  
  rsquared_values <- c(rsquared_values, rsquared)
}

tibble(rsquared = rsquared_values) %>% 
  ggplot(aes(x = rsquared)) +
  geom_histogram(bins = 30) +
  geom_boxplot(width = 5, position = position_nudge(y = -5)) +
  theme_minimal()
```

